---
title: "p8105_hw1_jw3638.Rmd"
output: github_document
date: "2025-09-19"
---
## Problem 1: 
##Load libraries
```{r }
library(tidyverse)
library(moderndive)
```

## Load dataset
```{r load_data}
data("early_january_weather")
```

The `early_january_weather` dataset contains:
`r nrow(early_january_weather)` rows and 
`r ncol(early_january_weather)` columns of hourly weather data collected at Newark Airport in January 2013.
The mean temperature during this time period was `r round(mean(early_january_weather$temp, na.rm = TRUE), 1)` °F.

The variables are: 
- `origin`: airport code (EWR)  
- `year`: year of observation  
- `month`: month of observation  
- `day`: day of month  
- `hour`: hour of day (0–23)  
- `time_hour`: date-time of the observation (`POSIXct`)  
- `temp`: temperature in °F  
- `dewp`: dew point in °F  
- `humid`: relative humidity (%)  
- `wind_dir`: wind direction (degrees from north)  
- `wind_speed`: wind speed (mph)  
- `wind_gust`: wind gust (mph)  
- `precip`: precipitation (inches)  
- `pressure`: air pressure (millibars)  
- `visib`: visibility (miles)  

##Making a scatterplot
```{r scatterplot}
weather_plot <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(
    title = "Temperature vs Time Colored by Humidity",
    x = "Time (Hours)",
    y = "Temperature (°F)",
    color = "Humidity (%)"
  )

weather_plot
```

##Plot analysis: 
Temperature changes in a repeating cycle-type pattern across time with highs and lows across the days.
There is a gradual increase in overall temperatures across the x-axis (January).
Humidity (indicated with different shades of blue) kind of moves in the opposite direction of temperature. Warmer conditions coordinate with less humidity.
Points group closely together and appear to be recorded at consistent and regular intervals.

##Export plot:
```{r save_plot}
ggsave("weather_scatterplot.png")
```

##Problem 2: step 1 --> Create data frame
```{r create_dataframe}
library(tidyverse)

## Set seed (for reproducibility)
set.seed(1234)

hw1_df <- tibble(
  vec_numeric = rnorm(10),                     
  vec_logical = vec_numeric > 0,                  
  vec_char = letters[1:10],                       
  vec_factor = factor(rep(c("low", "medium", "high"), length.out = 10)) 
)

hw1_df

##Taking the mean of each variable
mean(pull(hw1_df, vec_numeric))
mean(pull(hw1_df, vec_logical))
mean(pull(hw1_df, vec_char))     
mean(pull(hw1_df, vec_factor))   
```

##convert variables to numeric using as.numeric
```{r coercion, results='hide'}
as_numeric_logical <- as.numeric(pull(hw1_df, vec_logical))
as_numeric_char <- as.numeric(pull(hw1_df, vec_char))
as_numeric_factor <- as.numeric(pull(hw1_df, vec_factor))
```

##Mean results
When using as.numeric():
Logical: TRUE becomes 1, FALSE becomes 0 (works smoothly), but
Character: can't convert letters to numbers, and
Factor: converts to (1=low, 2=medium, 3=high)

